---
title: 'ANCOVA: Analysis of Covariance'
author: "Hotheifa Al-3Assi"
date: "`r Sys.Date()`"
output: pdf_document
---
\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{teach}
\end{figure}

\vspace{2cm}
Education is a vital part of our lives, helping us to understand everything and improve in all aspects. Therefore, developing the educational process and its methods is crucial.

This study aims to compare three teaching methods—traditional(Group1), online(Group2), and blended(Group3)—to determine if the teaching method significantly impacts students' final exam scores. Forty-Five students were randomly assigned to one of the three teaching methods. All students took a standardized final exam at the end of the course. Results from pre-tests, administered before the course began, were available and will be considered in the analysis.

Are there differences between teaching methods when controlling the pre-test? \ 

We want to know if there is a difference between the three teaching methods? \

Did the students' pre-test scores affect their final exam scores? \

Did the teaching methods and pre-test results influence the final exam?
\newpage

### The Data:
First six Rows:

```{r echo=FALSE}
mydata=read.csv("anxiety.csv",header = TRUE,sep=",") 
#######################################################################
library(knitr)
kable(mydata[1:6,-c(1)],col.names = c("Id","Group","Pre_test","Post_test"), align="l",caption ="The Data")

```

last six rows:
```{r echo=FALSE}
kable(mydata[39:45,-c(1,2)],col.names = c("Id","Group","Pre_test","Post_test"), align="l",caption ="The Data")
```


Dimensions of the Data set:

```{r echo=FALSE}
dim_table = data.frame(Dimension = c("Number of Rows", "Number of Columns"),Value = dim(mydata[,-c(1,2)]))
kable(dim_table,caption = "Dimensions of the Dataset")
```

Table 3 shows that the sample size is 45 and Variables we interest is 3: Group, Pre-test and Post-test.

### Varable type :
```{r echo=FALSE}
str_table <- data.frame(
  Variable = names(mydata[,-c(1,2)]),
  Type = sapply(mydata[,-c(1,2)], class),
  Example = sapply(mydata[,-c(1,2)], function(x) {vals <- x[!is.na(x)]    
vals <- head(vals, 3)
paste(as.character(vals), collapse = ",")}),row.names = NULL)
kable(str_table,caption = "Structure of the Dataset")
```
Table 4 shows that the variables pre-test,post-test is numerical and  The variable Group is Categorical. 

### Descriptives Statistics For Numirical Variables :
* The Numerical Descriptive Statistics:
```{r echo=FALSE}
library(psych)
des=describe(mydata[,-c(1,2,3)])
kable(des[,-c(1,2,6,7,10,11,12)],caption = "Descriptives Statistics for Numirical Variable",digits=2,longtable = FALSE)
```
Table 5 shows that the mean of pre-test is 16.92 and the Standard deviation is 1.49
and the mean of post-test is  15.24	 and Standard deviation 2.02.	


* The Graphical Descriptive Statistics:


```{r echo=FALSE,fig.align="center"}
boxplot(mydata[,-c(1,2,3)],col="red",horizontal = TRUE,main="Figure 1: Boxplot of Numirical Variables ")
```
Figure 1 shows that the mean of Pre-Test is higher than the Mean of the Post-Test
,while the spread of Post-test is higher than the spread of Pre-Test.


\newpage
### Categorical Variables :

Traditional Method (Group1), Online Method(Group2), and Blended Method(Group3)
All the Groups have the same size $n_i=15$ , $i=1,2,3$.


```{r echo=FALSE, fig.align='center', fig.height=3.5,fig.show='hold', out.width="90%"}
attach(mydata)
boxplot(pretest~group,col="red",main="Figure 2: Boxplot for Pre-Test with Groups")

```

Figure 2  shows that the mean for each group in the pre-test is almost equal, while there is a small difference in spread.

```{r echo=FALSE, fig.align='center',fig.height=3.5, fig.show='hold', out.width="90%"}
boxplot(posttest~group,col="red",main="Figure 3 : Boxplot for Post-Test with Groups")
```

Figure 3 shows that the mean for each group in the pre-test is difference , while there is a small difference in spread.

\newpage
## Inference Statistics:
For inferential statistics, we used regression analysis where the response variable was the Post-test which is numerical and the predictors were the pre-test which is numerical and the Groups which is Categorical.The results given in Table 6.

```{r echo=FALSE}
group1=factor(group)
library(knitr)
post_lm= lm( posttest~ pretest+ group1,data=mydata)
kable(as.data.frame(summary(post_lm)$coefficients) ,digits = 4,caption = " Regression Analysis")

```
### The results Show that : 
$\hat{B}_1$= 1.0278, When the pre-test increases by one Grade, the post-test average increases by 1.0278
With other factors remaining constant. 

$\hat{B}_2$=-0.6411, When moving from the traditional teaching method to online teaching methods will lower the average of post-test by -0.6411,With other factors remaining constant. 

$\hat{B}_3$=-2.9846, When moving from the traditional teaching method to blended teaching methods will lower the average of post-test by -2.9846,With other factors remaining constant. 

_____________________________________________________________________________________________________

### The Hypotheses:

Adjusted Post-Test: 

$H_0$:There are no statistically significant differences between the adjusted post-test means of the three teaching methods (after adjusting for pre-test) between the groups.

$H_1$:There is at least one statistically significant difference between the adjusted post-test means between the groups.

$p$-value for $\hat{\beta}_2$= 0.0001< $\alpha = 0.05$ , $p$-value for $\hat{\beta}_3$=0.0000 <$\alpha = 0.05$ ,  we reject $H_0$ at  $\alpha = 0.05$,There is statistically significant difference between the adjusted post-test means between the groups. 

### check the assumption :
We will now verify the assumptions of the linear regression model shown above. \
The results are shown in the figure:
```{r echo=FALSE,fig.align='center'}
par(mfrow=c(2,2))
plot(post_lm, which = 1, main = "Figure 4-A:")
plot(post_lm, which = 2, main = "Figure 4-B:")
plot(post_lm, which = 3, main = "Figure 4-c:")
plot(post_lm, which = 5, main = "Figure 4-D:")
```
* Figure 4-A: Residuals vs Fitted plot :
The points appear scattered around zero; there is no specific pattern, and linearity is achieved.\


* Figure 4-B: Q-Q Residuals plot : 
The points nearly lie on a straight line; the data follow a normal distribution. 

* Figure 4-C: Scale-Location :
The points appear scattered around zero; there is no specific pattern,Constant variance is achieved. 

* Figure 4-D: Residuals vs Leverage :
The diagnostic plot of residuals versus leverage indicated no highly influential observations, as no data points exceeded Cook’s Distance thresholds.


______________________________________________________________________________________________________
#### Difference between the three teaching methods: \

we used analysis of variance to find the difference between the three teaching methods,where the response is post-Test which is numerical and Groups which is categorical.
The results given in Table 7.
```{r echo=FALSE}
group_model = aov(posttest ~ group1, data = mydata)

anova_table = summary(group_model)[[1]]

kable(anova_table, digits = 5, caption = "ANOVA Table")

```
The Hypotheses: \
$H_0$:There is no statistically significant differences between means of the three teaching methods.\

$H_a$:There is statistically significant differences between means of the three teaching methods. \

$p-value$= 0.00002 < $\alpha = 0.05$ , we reject $H_0$ at $\alpha = 0.05$,There is statistically significant differences between means of the three teaching methods.
 
_________________________________________________________________________________________________

check the assumption :
```{r echo=FALSE, fig.align='center'}
par(mfrow=c(2,2))
plot(group_model, which = 1, main = "Figure 5-A:")
plot(group_model, which = 2, main = "Figure 5-B:")
plot(group_model, which = 3, main = "Figure 5-c:")
plot(group_model, which = 5, main = "Figure 5-D:")
```
* Figure 5-A: Residuals vs Fitted plot :
The points appear scattered around zero; there is no specific pattern.\


* Figure 5-B: Q-Q Residuals plot : 
The points nearly lie on a straight line; the data follow a normal distribution. 

* Figure 5-C: Scale-Location :
The points appear scattered around zero; there is no specific pattern,Constant variance is achieved. 

* Figure 5-D: Residuals vs Leverage :
The diagnostic plot of residuals versus leverage indicated no highly influential observations, as no data points exceeded Cook’s Distance thresholds.

\newpage
### Pairwise Comparisons:
Since we rejected the null hypothesis, this means that there are differences between the means. The comparison method was used Bonferroni. Table 8 shows the results:
```{r echo=FALSE}
com=pairwise.t.test(posttest,group1,"bonferroni")
p_matrix <- com$p.value
kable(p_matrix, digits = 4, caption = "Pairwise Comparisons (Bonferroni Adjustment)")
```
The Hypotheses: \

* Traditional Method (Group1) and Online Method(Group2).\

$H_0: \mu_{Traditional} = \mu_{Online}$.\

$H_1: \mu_{Traditional} \neq \mu_{Online}$.\

$p-value$=0.2074 < $\alpha = 0.05$, we do not reject $H_0$ at $\alpha = 0.05$,there is no difference between Traditional and Online Methods.\

________________________________________________________________________________________________
* Traditional Method (Group1) and Blended Method(Group3).\ 

$H_0: \mu_{Traditional} = \mu_{Blended}$.\

$H_1: \mu_{Traditional} \neq \mu_{Blended}$.\

$p-value$=	0.0000 < $\alpha = 0.05$, we reject $H_0$ at $\alpha = 0.05$,there is difference between Traditional and Blended Methods.

________________________________________________________________________________________________
* Online Method (Group2) and Blended Method(Group3).\

$H_0: \mu_{Online} = \mu_{Blended}$.\

$H_1: \mu_{Online} \neq \mu_{Blended}$.\

$p-value$=	0.0051 < $\alpha = 0.05$, we reject $H_0$ at $\alpha = 0.05$,we reject $H_0$ at $\alpha = 0.05$,there is difference between Online and Blended Methods.



\newpage

Pre-Test on Post-Test: \

To Study the relationship between Pre-Test and Post-Test  we used regression analysis where the response variable was the Post-test which is numerical and the predictors were the pre-test which is numerical. The results given in Table 9.
```{r echo=FALSE}
pree_model=lm(posttest~pretest,data=mydata)
kable(as.data.frame(summary(pree_model)$coefficients) ,digits = 4,caption = " Regression Analysis")
```
#### The results Show that :\


$\hat{B}_1$= 1.0094	, When the pre-test increases by one Grade, the post-test average increases by 1.0278.\

__________________________________________________________________________________________________

The Hypotheses: \

$H_0:$:There is no statistically significant effect of pre-test on post-test.
$H_1:$There is  statistically significant effect of pre-test on post-test.\

$p$-value for $\hat{\beta}_2$= 0.0000< $\alpha = 0.05$ ,we reject $H_0$ at  $\alpha = 0.05$,there is  statistically significant effect of pre-test on post-test.\

\newpage
check the assumption :

```{r echo=FALSE, fig.align='center'}

par(mfrow=c(2,2))
plot(pree_model, which = 1, main = "Figure 6-A:")
plot(pree_model, which = 2, main = "Figure 6-B:")
plot(pree_model, which = 3, main = "Figure 6-c:")
plot(pree_model, which = 5, main = "Figure 6-D:")
```
* Figure 6-A: Residuals vs Fitted plot :
The points appear scattered around zero; there is no specific pattern, and linearity is achieved.\


* Figure 6-B: Q-Q Residuals plot : 
The points nearly lie on a straight line; the data follow a normal distribution. 

* Figure 6-C: Scale-Location :
The points appear scattered around zero; there is no specific pattern,Constant variance is achieved. 

* Figure 4-D: Residuals vs Leverage :
The diagnostic plot of residuals versus leverage indicated no highly influential observations, as no data points exceeded Cook’s Distance thresholds.

\newpage
### Interaction:

For Interaction we used regression analysis where the response variable was the Post-test which is numerical and the predictors were the pre-test which is numerical and the Groups which is Categorical.The results given in Table 10.
```{r echo=FALSE}
group1=factor(group)
library(knitr)
post_lm1= lm( posttest~ pretest+ group1+ pretest:group1,data=mydata)
kable(as.data.frame(summary(post_lm1)$coefficients) ,digits = 4,caption = "Regression Analysis With Interaction")
```
#### The results Show that :\

$\hat{B}_1$= 1.0128, When the pre-test increases by one unit, the post-test average increases by 1.0128
With other factors remaining constant. \

$\hat{B}_2$=-1.3770, when Using online teaching methods will decreases the average of post-test by -1.3770
With other factors remaining constant. \

$\hat{B}_3$=-2.9349	,when Using blended teaching methods will decreases the average of post-test by -2.9349
With other factors remaining constant. 

$\hat{B}_4$= 0.0438,when the interaction between the pre-test and the group2 increases by one unit   the Post-test average increase by 0.0438 With other factors remaining constant.

$\hat{B}_5$=-0.0030	when the interaction between the pre-test and the group3 increases by one unit   the Post-test average decreases by -0.0030 With other factors remaining constant. 

### The Hypotheses:

Interaction:

$H_0$:There is no interaction effect between pre-test scores and group 2 membership on post-test scores. 

$H_1$:There is an interaction effect between pre-test scores and group 2 membership on post-test scores.  

p$-value for $\hat{\beta}_2$= 0.6623 > $\alpha = 0.05$ , we Do not reject $H_0$ at  $\alpha = 0.05$,
There is no interaction effect between pre-test scores and group 2 membership on post-test scores. 

_____________________________________________________________________________________________________


$H_0$:There is no interaction effect between pre-test scores and group 3 membership on post-test scores. 

$H_1$:There is an interaction effect between pre-test scores and group 3 membership on post-test scores.


p$-value for $\hat{\beta}_2$= 0.9784 > $\alpha = 0.05$ , we Do not reject $H_0$ at  $\alpha = 0.05$,
There is no interaction effect between pre-test scores and group 3 membership on post-test scores. 

______________________________________________________________________________________________________


### Compare models: 
we use Anova to compare the two models 

```{r echo=FALSE}
kable(anova(post_lm,post_lm1))
```


$H_0$: $B_4$=$B_5$=0. \

$H_a$: $B_4$$\neq$$B_5$$\neq 0$

$p-value$ = 0.8815 > $\alpha = 0.05$, we do not reject $H_0$ at $\alpha = 0.05$,
The models are equal or The intersection has no effect or importance.































